\begin{thebibliography}{9}

\bibitem{go}
D. Silver et al., "Mastering the game of Go with deep neural networks and tree search",
\textit{Nature}, 529, p. 484–489, 2016.

\bibitem{foundations}
F. Cucker, S. Smale, "On the mathematical foundations of learning",
\textit{Bull. Amer. Math. Soc.}, 39, p. 1-49, 2002.

\bibitem{hist1}
W. Krauth, M. Mezard, "Machine Learning algorithms with optimal stability in neural networks",
\textit{Journal of Physics A: Mathematical and General}, 20, L745–L752, 1987.

\bibitem{hist2}
J. J. Hopfield, "Neural networks and physical systems with emergent collective computational abilities"\textit{Proceedings of the National Academy of Sciences, 79}, 8, p. 2554–2558, 1982.

\bibitem{tap_train}
M. Gabri\'e, E. W. Tramel, F. Krzakala,
Training Restricted Boltzmann Machines via the Thouless-Anderson-Palmer Free Energy,
\textit{Advances in Neural Information Processing Systems (NIPS)}, 28, pages 640--648, 2015.

\bibitem{tap}
E. W. Tramel, M. Gabri\'e, A. Manoel, F. Caltagirone, F. Krzakala, \textit{arXiv:1702.03260}

\bibitem{monasson}
J. Tubiana, R. Monasson, "Emergence of Compositional Representations in Restricted Boltzmann Machines",
\textit{Phys. Rev. Lett. 118, 138301}, 2017.

\bibitem{Hinton_CD}
G. E. Hinton, “Training products of experts by minimizing Contrastive divergence,”
\textit{Neural computation}
, vol. 14, pp. 1771-1800, 2002.

\bibitem{PCD}
T. Tieleman, “Training restricted Boltzmann machines using approximations to the likelihood gradient,”
\textit{ICML}, Vol. 307, p. 7, 2008.

\bibitem{Hinton_guide}
G. E. Hinton, "A Practical Guide to Training Restricted Boltzmann Machines", \textit{Proceedings of Neural Networks: Tricks of the Trade (2nd ed.)}, pages 599-619, 2012. 

\bibitem{PCA}
I.T. Jolliffe, "Principal Component Analysis",
\textit{Springer-Verlag New York}, 2002.

\bibitem{SK}
D. Sherrington, S. Kirkpatrick,
"Solvable Model of a Spin-Glass",
\textit{Phys. Rev. Lett.}, Vol. 35, 1975.

\bibitem{ht_exp}
A. Georges, J. S. Yedidia,
"How to expand around mean-field theory using high-temperature expansions",
\textit{Journal of Physics A: Mathematical and General}, Volume 24, Number 9, 1991

\bibitem{TAP}
D. J. Thouless , P. W. Anderson, R. G. Palmer,
"Solution of 'Solvable model of a spin glass'",
\textit{Philosophical Magazine}, Vol. 35:3, p. 593-601, 1977

\bibitem{conv}
E. Bolthausen, "An Iterative Construction of Solutions of the TAP Equations for the Sherrington–Kirkpatrick Model", \textit{Commun. Math. Phys.}, Vol. 325, pages 333-366, 2014.

\bibitem{mnist}
http://yann.lecun.com/exdb/mnist/

\bibitem{gibbs}
G. Casella and E. I. George, "Explaining the Gibbs Sampler",
\textit{The American Statistician},
Vol. 46, No. 3, pp. 167-174, 1992.

\bibitem{leroy}
C. Leroy made an investigation on TAP fixed points and free energy landscape while at Laboratoire de Recherche en Informatique, working in my same team.

\bibitem{Nishimori}
H. Nishimori, "Statistical Physics of Spin Glasses and Information Processing: An Introduction",
\textit{Oxford University Press}, 2001

\bibitem{MP_law}
Marchenko, V.A. and Pastur, L.A., "Distribution of Eigenvalues for Some Sets of Random Matrices", \textit{Sbornik: Mathematics}, Vol. 1, pages 457-483, 1967.

\bibitem{ganguli}
A. M. Saxe, J. L. McClelland, S. Ganguli, "Exact solutions to the nonlinear dynamics of learning in deep linear neural networks",
\textit{arXiv:1312.6120}, 2014.

\end{thebibliography}
